---
title: "Unit 11 Assignment"
date: "November 22, 2016"
author: "Angela Horacek, James Logan Gage, Trace Smith"
knit: (function(inputFile, encoding) {
      rmarkdown::render(inputFile, 
                        encoding=encoding, output_file='Papers/Report.html') } )

output: 
  html_document:
    keep_md: true
---
# Introduction

- Data set "elecequip" shows the number of new orders for electrical equipment (computer, electronic and optical products) in the Euro area (16 countries). The data for this assignment is loaded from the 'Forecasting: principles and practice' (i.e. fpp) library. The source for the data 'hsales' which is the monthly sales of new one-family houses sold in the USA since 1973 is from: "Makridakis, Wheelwright and Hyndman (1998) Forecasting: methods and applications, John Wiley & Sons: New York. Chapter 3."

```{r, include=FALSE}
# This code sets up the environment by running the Analysis/SetupEnvironment.R script. Output is not included in the RMarkdown output.
knitr::opts_chunk$set(echo = TRUE)
source("Analysis/SetupEnvironment.R")
```

#### a) Plot the time series. Can you identify seasonal fluctuations and/or a trend?

```{r}
source("Analysis/hsalesPlot.R", echo=TRUE)
```

- There appears to be a seasonality to the data. The data seem to be oscillating up and down corresponding to the calender year. In addition, a trend in the form of a sinusodal wave appears with a trough.

#### b) Use a classical decomposition to calculate the trend-cycle and seasonal indices.

```{r}
source("Analysis/hsalesDecomp.R", echo=TRUE)
```

- The seasonal indice look to cycle every calendar year and the trend indice is close to a 10 year cycle.

#### c) Do the results support the graphical interpretation from part (a)?

- The results from the decomposition graphics indicate there is not only a seasonal component but also a trend component bolstering the interpretation of the original graph.

#### d) Compute and plot the seasonally adjusted data.
 
- Decompose a time series into seasonal, trend and irregular components using:
 
```{r}
source("Analysis/hsalesSeasonal.R", echo=TRUE)
```

#### e) Change one observation to be an outlier (e.g., add 500 to one observation), and recompute the seasonally adjusted data. What is the effect of the outlier?

```{r}
source("Analysis/hsalesChangeOne.R", echo=TRUE)
```

- The outlier observation in the middle of the data increases the magnitude of the the seasonally adjusted data not only at that particular time but also on both sides of that time. Similarly, the trend line also peaks at that year pulling the trend curve up around that time distorting the sinusodial wave pattern.

#### f) Does it make any difference if the outlier is near the end rather than in the middle of the time series?

```{r}
source("Analysis/hsalesOutlierEnd.R", echo=TRUE)
```

- The outlier observation near the end of the data increases the magnitude of the the seasonally adjusted data not only at that particular time but also on both sides of that time; This is the same effect of the outlier in the middle except at a different time. Similarly, the trend line also peaks at that year pulling the trend curve up around that time distorting the sinusodial wave pattern.

#### g) Now use STL to decompose the series.

```{r}
source("Analysis/hsalesOutlierSTL.R", echo=TRUE)
```


